Comparing the Jacobi Method and LLL lattice reduction algorithms for cryptographic applications

### 2014-12-08 report/paper.tex
Lagrange + Jacobi Variant

- lattice prime determinant p 
- unique HNF form
- 1 on the other places of the diagonal
- all entries under the p are smaller than P in absolute value.

--->

TODO: Still unsure about how to phrase it

ressources from Nicolas:
A few (understandable) papers on random lattices:

- D. Goldstein and A. Mayer. On the equidistribution of Hecke points.
Forum Math., 15(2):165–189, 2003.

basically, proves that random HNFs converge to the Haar Measure

- Random lattices and a conjectured 0 -1 law about their polynomial time
computable properties, Ajtai, FOCS 2002

gives the interpretation that for any property verifiable in polynomial
time,
if the property is asymptotically true with x percent on the HNF random
integer lattices when the dimension grow,
then the property is true with x percent on the Haar measure.

@@

\section{The Fast Jacobi Method}
\subsection{Lagrange-reduced basis}

@@

\subsection{Lagrange algorithm}

Reducing

Complexity quadratic

\subsection{Our implementation}

>>>>

We say that a lattice is a random lattice $L$ if under HNF form its basis matrix $\vec{B}$ has the following properties:

\begin{itemize}
\item the diagonal has 1 for all it's entries except a prime number $P$. Hence, the $\det(\vec{B})$ is prime.
\item All entries of the matrix under $P$ are smaller than $P$ in absolute value.
\end{itemize}

@@

\section{The Jacobi Method}

In 2012, Sanzheng Qiao suggested to use a Jacobi method for lattice basis reduction\cite{originalJacobiMethodLatticeBasisReduction}. The idea of the Jacobi method is to use the reduce larger dimension lattice basis by using the Lagrange algorithm iteratively.

\subsection{Lagrange-reduced basis}

@@

\subsection{Lagrange algorithm}
\label{sec:lagrangeAlgorithm}

Lagrange's algorithm solves the lattice reduction problem exactly for lattice basis of dimension 2. It is a variant of Euler's centered algorithm that is used to compute the greatest common divider of two numbers. GCD can be thought to be a one dimension reduction problem. The GCD is the first minimum of the lattice $n \mathbb{Z} + m \mathbb{Z}$ spanned by $n$ and $m$.

\begin{algorithm}[H]
\caption{Lagrange algorithm}
\label{lagrangeAlgorithm}
\begin{algorithmic}[1]
\REQUIRE a basis ($\vec{b_1, b_2}$) of a lattice $L$ of dimension 2.
\ENSURE a Lagrange reduced reduced basis $(\vec{b_1, b_2})$

\IF{$\|\vec{b_1} \| < \| \vec{b_2} \| $} \STATE{swap $\vec{b_1}$ and $\vec{b_2}$} \ENDIF
\REPEAT \STATE{
$q = \lfloor \frac{ \langle \vec{b_1} \vec{b_2} \rangle }{\| v\|^2} \rceil $\\
$r \leftarrow \vec{b_1} - q \vec{b_2}$\\
$\vec{b_1} \leftarrow \vec{b_2}$\\
$\vec{b_2} \leftarrow r$
} \UNTIL{$\| u\| \leq \| v \| $}

\subsection{The Generic Jacobi Method}
We call \emph{generic Jacobi method} the algorithm that simply applies the Lagrange algorithm to every pair of the basis vectors for an $n$ dimensional lattice. This method was introduced\cite{originalJacobiMethodLatticeBasisReduction} by Qiao in 2012 claiming experimental results to be considerably faster than LLL. We haven't been able to reproduce the claimed running times of this algorithm.

From the definition of a Lagrange-reduced basis, we can define a Jacobi-reduced basis such as:
\begin{align}
\langle \vec{b_i} \vec{b_j} \rangle &\leq \frac{1}{2} \| \vec{b_i}\|^2 \text{ (for all $1 \leq i < j \leq n$)}.\\
\| \vec{b_i}\| &\leq \|\vec{b_j} \| \text{ (for all $1 \leq i < j \leq n$)}
\end{align}

\begin{algorithm}[H]
\caption{Generic Jacobi Method}
\label{genericJacobiMethod}
\begin{algorithmic}
\REQUIRE a basis matrix ($\vec{b_1,...,b_n}$)
\ENSURE a generic-Jacobi reduced basis ($\vec{b_1,...,b_n}$)

\WHILE{not all pairs ($\vec{b_i,b_j}$) satisfy both generic-Jacobi reduction conditions}
    \FOR{$i = 1$ \TO $n-1$ }
        \FOR{$j = i + 1$ \TO $n$ }
            \STATE {$[\vec{b_i, b_j}]$} = \emph{Lagrange$(\vec{b_i, b_j})$}
        \ENDFOR
    \ENDFOR
\ENDWHILE

\subsection{A Jacobi Variant}

In this part, we present the Jacobi-variant that was used for running our benchmarking.

\subsubsection{$\omega$-Lagrange-reduced conditions}
An $\omega$ is introduced to 
There are two conditions for a basis to be $\omega$-Lagrange-reduced.
\[
\begin{cases}
\lfloor \vec{G}_{ij} / \vec{G}_{ss} \rceil \leq 1, \\
\omega^2 \vec{G}_{ll} \leq \vec{G}_{ii} + \vec{G}_{jj} - 2|\vec{G}_{ij}|

\end{cases}
\]  where $1/\sqrt{3} \leq \omega < 1$.

\subsubsection{Size-reduction Lagrange}
The size-reduction algorithm is a variant of the Lagrange algorithm that we presented in section \ref{sec:lagrangeAlgorithm}. The reduction is only performed if the $\omega$-Lagrange-reduced conditions are not satisfied. In which case, both the unimodular reduction matrix $\vec{Z}$ and the Gram matrix $\vec{G}$ are updated by applying the Lagrange reduction to the longest of the basis vectors at index $i$ or $j$.

\begin{algorithm}[H]
\caption{LagrangeIT}
\label{customLagrangeIT}
\begin{algorithmic}
\REQUIRE The matrices $\vec{G, Z}$, a pair of indices $(i,j):i<j$ and a parameter $\omega$
\ENSURE  Updated $\vec{G, Z}$ where one Lagrange iteration was performed on the $i$th and $j$th basis vectors.

\STATE {$s \leftarrow i$}
\STATE {$l \leftarrow j$}

\IF{$\vec{G}_{ii} > \vec{G}_{jj}$}
    \STATE {$s \leftarrow j; l \leftarrow i$}
\ENDIF

\STATE{$q \leftarrow \lfloor \frac {\vec{G}_{ii}}{\vec{G}_{ss}} \rceil $}

\IF {Verify both $\omega$-Lagrange-reduced conditions}

\STATE{$\vec{Z_l} -= q* Z_s $}
\STATE{$\vec{G_l} -= q* G_s $}

\FOR{every column $k$ of $\vec{G}$}
\STATE{$\vec{G}_{lk} = \vec{G}_{kl} $}
\ENDFOR

\STATE{$\vec{G}_{ll} -= q* \vec{G}_{ls} $}
\ENDIF

\subsection{Jacobi-variant algorithm}
This Jacobi-variant performs Lagrange reductions iteratively on the Gram matrix until no more reduction is done. That is when both $\omega$-reduced Lagrange conditions are satisfied.

\begin{algorithm}[H]
\caption{Jacobi-Variant Reduction}
\label{jacobiVariantReduction}
\begin{algorithmic}
\REQUIRE a basis matrix ($\vec{B = b_1,...,b_n}$) and $\omega$
\ENSURE a reduced basis ($\vec{b_1,...,b_n}$) where each pair of vectors is $\omega$-Lagrange reduced
\STATE{$\vec{G} = \langle \vec{B} \rangle$, $\vec{Z} = I_n$}
\WHILE{LagrangeIT method reduced the basis vectors}
    \FOR{$i = 1$ \TO $n-1$ }
        \FOR{$j = i + 1$ \TO $n$ }
            \STATE {$[\vec{G,Z}]$} = \emph{LagrangeIT$(\vec{G,Z},i,j,\omega)$}
        \ENDFOR
    \ENDFOR
\ENDWHILE

\subsection{Our Implementation}

We implemented this Jacobi algorithm in C++ with the \emph{newNTL} library. The FPLLL implementation was used for LLL reductions.

\subsection{Experimental Results}

We notice that the Jacobi algorithm does reduce the bases further than LLL. This comes that the LLL reduction definition is more.
The Jacobi method requires the vectors to be

We show two examples of bases that are LLL-reduced but not $\omega > 0.9$ Jacobi reduced.
\begin{enumerate}

\item \[
\vec{B} = \begin{bmatrix}
  \vec{b_1} \\
  \vec{b_2} \\
  \vec{b_3}
 \end{bmatrix} = \begin{bmatrix}
  0 & 2 & 0 \\
  0 & 1 & 2 \\
  2 & 0 & 0
 \end{bmatrix}
\]

\item \[
\vec{B} = \begin{bmatrix}
  \vec{b_1} \\
  \vec{b_2} \\
  \vec{b_3}
 \end{bmatrix} = \begin{bmatrix}
  4/3 & 0 & 0 \\
  2/3 & \sqrt{4/3} & 0 \\
  2/3 & \sqrt{4/3}/2 & 1
 \end{bmatrix}
\]
The Jacobi Method will find a shortest vector by subtracting $\vec{b_1}-\vec{b_2}$, while LLL won't be able to reduce further.
\end{enumerate}
